Omics-based Single Step Trait Prediction

### 2017-03-30 paper/manuscript.tex
Fix typos in manuscript

\subsection{Maize hybrids}
Hybrid breeding --- with noteworthy commercial applications in pig
\cite{Xiang2016,Tusell2016} and maize breeding \cite{TheRoyalSociety2009} --- 
is a particularly challenging application field for prediction tools 
\cite{Kadam2016}.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology \cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.
\citeNP{Westhues2017} have shown that, in hybrid breeding, the probability 
of successfully selecting the best observed genotypes based on the best 
predicted candidates is a strongly convex function of the predictive ability.
Thus, even the minor gains when using transcriptomic and genomic data in a
single step might justify additional investments in RNA-seq at least of a
subset of genotypes that represents the genetic space of the breeding 
material well.
For the hybrid data, the predictive abilities of genomic information
for the whole set of available parent lines rose considerably for several
agronomic traits.
Hence, the inclusion of trancsriptomic data could not further improve upon
this result.
Predictive abilities obtained when using pedigree data could be 
improved considerably when, depending on the agronomic trait, combined with 
either transcriptomic or genomic information.
Evaluations of core sets for the parental inbred lines of the maize hybrids
used throughout this study were outside the scope of this paper, but
results on the inbred lines have shown that homogeneous coverage of the
genetic space by the incomplete predictor can considerably leverage the
predictive ability in single step prediction compared to using only the
complete predictor.

\Genetics2level{Coverage of the genetic space}
Animal populations are so large that it the assembly of a core set of
individuals with records on multiple predictors can effectively be done at 
random once more than 10,000 animals have pedigree records 
\cite{Fragomeni2015,Lourenco2015,Masuda2016}, but in plant breeding programs,
pedigrees are typically considerably smaller.
Single step prediction offers the promise of leveraging the predictive ability
for all genotypes by borrowing the superior performance of an auxiliary
predictor on only a subset of genotypes, thereby reducing predictor costs.
Hence, we were interested in determining which individuals should be 
complemented with information on another predictor.
When using core sets of varying sizes, which were established with the
constraint of maximizing the average genetic distance among their members, we
noticed strong fluctuations in predictive abilities when moving from small core
sets to large core sets.
For traits that were better predicted by transcriptomic than by genomic
information in the core set, predictive abilities typically peaked around
core set size of 0.5/0.6, then dropped sharply at 0.7 and started to increase
again with a size of 0.9.
A principal component analysis showed that, at a core set size of 0.7, 
genotypes assigned to the A3 ancestral population (Fig. \ref{fig:PopStructure})
were largely elements of the core set whereas genotypes with a large share at 
the A1 ancestral population were predominantly excluded from the core set.
At a core set size of 90\% of the original size, the majority of the A1
genotypes were finally covered by the core set as well.
This result suggests a negative bias in the imputation of genotypes from the A1
population and could explain the drop in predictive ability at a core set size
of 0.7, which is in agreement with the reduced predictive abilities observed
when stripping A1 genotypes from their transcriptomic information (Fig.
\ref{fig:InbredResults}B).

>>>>

\subsection{Maize hybrids}
Hybrid breeding --- with noteworthy commercial applications in pig
\cite{Xiang2016,Tusell2016} and maize breeding \cite{TheRoyalSociety2009} --- 
is a particularly challenging application field for prediction tools 
\cite{Kadam2016}.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology \cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.
\citeNP{Westhues2017} have shown that, in hybrid breeding, the probability 
of successfully selecting the best observed genotypes based on the best 
predicted candidates is a strongly convex function of the predictive ability.
Thus, even the minor gains when using transcriptomic and genomic data in a
single step might justify additional investments in RNA-seq at least of a
subset of genotypes that represents the genetic space of the breeding 
material well.
For the hybrid data, the predictive abilities of genomic information
for the whole set of available parent lines rose considerably for several
agronomic traits.
Hence, the inclusion of trancsriptomic data could not further improve upon
this result.
Predictive abilities obtained when using pedigree data could be 
improved considerably when, depending on the agronomic trait, combined with 
either transcriptomic or genomic information.
Evaluations of core sets for the parental inbred lines of the maize hybrids were 
outside the scope of this study, but results on the inbred lines have shown that 
homogeneous coverage of the genetic space by the incomplete predictor can 
considerably leverage the predictive ability in single step prediction compared 
to using only the complete predictor.

\Genetics2level{Coverage of the genetic space}
Whereas animal breeding populations are oftentimes so large that the assembly 
of a core set of individuals with records on multiple predictors can effectively 
be done at random once more than 10,000 animals have pedigree records 
\cite{Fragomeni2015,Lourenco2015,Masuda2016}, pedigrees in plant breeding 
programs are typically considerably smaller.
Single step prediction offers the promise of leveraging the predictive ability
for all genotypes by borrowing the superior performance of an auxiliary
predictor on only a subset of genotypes, thereby reducing predictor costs.
Hence, we were interested in determining which individuals should be 
complemented with information on another predictor.
When using core sets of varying sizes, which were established with the
constraint of maximizing the average genetic distance among their members, we
noticed strong fluctuations in predictive abilities when moving from small core
sets to large core sets.
For traits that were better predicted by transcriptomic than by genomic
information in the core set, predictive abilities typically peaked around
core set size of 0.5/0.6, then dropped sharply at 0.7 and started to increase
again with a size of 0.9.
A principal component analysis showed that, at a core set size of 0.7, 
genotypes assigned to the A3 ancestral population (Fig. \ref{fig:PopStructure})
were largely elements of the core set whereas genotypes with a large share at 
the A1 ancestral population were predominantly excluded from the core set.
At a core set size of 90\% of the original size, the majority of the A1
genotypes were finally covered by the core set as well.
This result suggests a negative bias in the imputation of genotypes from the A1
population and could explain the drop in predictive ability at a core set size
of 0.7, which is in agreement with the reduced predictive abilities observed
when stripping A1 genotypes from their transcriptomic information (Fig.
\ref{fig:InbredResults}B).
